---
title: "Advanced Statistical Methods Hw7"
author: "Do Hyup Shin"
date: "`r format(Sys.Date())`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 9.3
Redraw Figure 9.2, changing the “knot” location from 11 to 12.


### Solution
First, we will check case Arm_A of the NCOG. In the book, n = number at risk(patients), y = number of deaths, l = lost to followup, h = hazard rate y/n; $\hat{S}$ = life table survival esimate.
Next, we assume that $y_i \overset{indep}{\sim} B(n_i, h_i) \quad \forall i = 1, 2, \dots,47$. We will use a generalized linear model.
Let $\mu_i = E(y_i) = n_ih_i \quad \forall i = 1, 2, \dots , 47$.
The pdf of $y_i$ is 

\begin{align*}
  f(y_i) & = \binom{n_i}{y_i}h_i^{y_i}(1-h_i)^{n_i-y_i} = exp(y_i log\frac{h_i}{1-h_i} + n_i log(1 - h_i) + log\binom{n_i}{y_i})\\
  & = exp(y_i log\frac{\mu_i}{n_i-\mu_i} + n_i log(\frac{n_i - \mu_i}{n_i}) + log\binom{n_i}{y_i}) \\
  & =  exp(y_i\theta_i - n_i log(1 + e^{\theta_i}) + log\binom{n_i}{y_i}) \quad \text{ where } \theta_i = log\frac{\mu_i}{n_i - \mu_i}
\end{align*}

Define $b(\theta_i) = n_ilog(1+e^{\theta_i}), \quad a(\phi) = 1$, $\quad g(\mu_i) = \theta_i = log\frac{\mu_i}{n_i - \mu_i} =  \lambda_i = x_i^T\alpha$ where $x_i = (x_{i1}, x_{i2}, x_{i3}, x_{i4}) =(1, i, (i - 12)^2\_ ,(i-12)^3\_ )^T, \quad \alpha = (\alpha_1,\alpha_2,\alpha_3,\alpha_4)^T$ and we can express $\mu_i = n_i \cfrac{e^{\theta_i}}{1+e^{\theta_i}}$. By using this form, calculate log likelihood function $l(\alpha)$.

\begin{align*}
  l(\alpha) & = \sum_{i = 1}^{47}(y_i\theta_i - b(\theta_i) + log\binom{n_i}{y_i}) \\
   & = \sum_{i=1}^{47}(y_i x_i^T \alpha - n_i log(1+e^{x_i^T \alpha}) + log\binom{n_i}{y_i})
\end{align*}


Let $S(\alpha)$ be score function. 
\begin{align*}
  S(\alpha_j) & = \frac{\partial S}{\partial \alpha_j} \\
            & = \sum_{i=1}^{47} ( y_i x_{ij} - n_{i}\cfrac{e^{x_i ^T \alpha}}{1 + e^{x_i ^T \alpha}} x_{ij} ) \\
            & = \sum_{i=1}^{47}(y_i - \mu_i)x_{ij} = 0 \quad \forall j = 1, 2, 3, 4
\end{align*}

We will find the mle of $\alpha_j$ satisfying $S(\alpha_j) = 0$.
But, we cannot find the exact solution. So we will use the Iteratively Rewighted least square algorithm method.

Let $g(y_i) \approx z_i = g(\mu_i) + g'(\mu_i)(y_i - \mu_i) = \lambda_i + \frac{d\lambda_i}{d\mu_i}(y_i - \mu_i)$ with $var(z_i) = \cfrac{var(y_i)}{(\frac{d\mu_i}{d\lambda_i})^2}$.
Then, $g'(\mu_i) = \frac{n_i}{\mu_i(n_i - \mu_i)}$ and define
$D = diag(\frac{\partial\mu_i}{\partial\lambda_i}) = diag(\frac{\mu_i(n_i - \mu_i)}{n_i})$ and $V = diag(var(y_i)) = diag(\mu_i(\frac{n_i - \mu_i}{n_i}))$.

The IRLS algorithm works as follows.
\begin{enumerate}
  \item[Step1] Define the kth vector value $\alpha_{(k)}$.
  \item[Step2] By using $\alpha_{(k)}$, Calculate $\lambda_{(k)}, \mu_{(k)}, z_{(k)},D_{(k)}, V_{(k)}$.
  \item[Step3] Calculate $\alpha_{(k+1)} = (X^TW_{(k)}X)^{-1}X^TW_{(k)}z_{(k)}$.
  \item[Step4] Continue above process until the $\alpha_{(k)}$ converges.
  
\end{enumerate}

Actually, we can simplify $W = DV^{-1}D = diag(\frac{\mu_i(n_i - \mu_i)}{n_i}) diag(\frac{n_i}{\mu_i(n_i - \mu_i)})diag(\frac{\mu_i(n_i - \mu_i)}{n_i}) = diag(\frac{\mu_i(n_i - \mu_i)}{n_i}) = D$.
So, we can find the numerical solution of the mle $\alpha$.

```{r}
library(dplyr)
library(ggplot2)
NCOG_data <- read.csv("https://web.stanford.edu/~hastie/CASI_files/DATA/ncog.txt", sep = " ")

#Separate A and B
Arm_A <- subset(NCOG_data,subset = arm == "A")
Arm_B <- subset(NCOG_data,subset = arm == "B")

## harzard function of Arm_A
num_A = nrow(Arm_A)
Km_A <- matrix(0, 47, 4)
copy_A = Arm_A

#make the 47 * 4 matrix
for(i in 1:47){
  Km_A[i,1] = nrow(copy_A)
  inter_A = copy_A %>% filter(t > 30.4*(i-1) & t <= 30.4*i)
  Km_A[i, 2] <- length(which(inter_A$d == 1))
  Km_A[i, 3] <- length(which(inter_A$d == 0))
  Km_A[i, 4] <- Km_A[i, 2] / Km_A[i,1]
  copy_A = copy_A %>% filter(t > 30.4*i)
}

x_a <- matrix(0, 47, 4)
for(i in 1:47){
  if(i <= 11){
    x_a[i,] <- c(1, i, (i-12)^2, (i-12)^3)
  }
  else{
    x_a[i,1:2] <- c(1, i)
  }
}

#number of patients
n_a = Km_A[,1]

#number of deaths
y_a = Km_A[,2]

## Iteratively Rewighted least square(IRLS)
# initial value of alpha
alpha_0 = c(-1, -0.01, 0.1, 0.01)
alpha = matrix(0, 1001, 4)
alpha[1,] = alpha_0

#iterate 1000 times
for(i in 1:1000){
  
  lambda = x_a %*% alpha[i,]
  
  # mu_k = n * exp(lambda_k)/ (1 + exp(lambda_k))
  mu = n_a * exp(lambda) / (1 + exp(lambda))
  
  # z_k = lambda_k + n / (mu_k * (n - mu_k)) * (y - mu_k)
  z = lambda  + (n_a / (mu * (n_a - mu)) ) * (y_a - mu)
  
  
  # D_k = diag(mu_k(n - mu_k) / n)
  D = diag(as.numeric(mu*(n_a - mu)/n_a))
  
  
  # V_k = diag(mu_k (1 - mu_k/n))
  V = diag(as.numeric(mu* (1 - mu/n_a)))
  # W_k = D %*% solve(V) %*% D
  
  W = D
  
  # alpha_(k+1) = (X^t W_k X)^(-1) X^t W_k Z_k
  alpha[i+1, ] = solve((t(x_a) %*% W %*% x_a)) %*% t(x_a) %*% W %*% z
}
alpha_hat_a = alpha[1001,]

#The numerical solution of mle
alpha_hat_a
```


Thus, the numerical mle of $\alpha$ in Arm_A case is $\hat{\alpha} = (-2.8303, -0.01587,  0.08267,  0.0085)^T$. Next, we'll draw the graph of harzard ratio.


```{r}
#harzard ratio
harzard_func <- function(x){
  h = 1/(1 + exp(-x%*%alpha_hat_a))
  return(h)
}

#cubic linear spline function
spline_func <- function(x){
  if(x<= 12){
    return(c(1, x, (x-12)^2,(x-12)^3))
  }
  else{
    return(c(1, x, 0, 0))
  }
}

#range of x
range_xa = seq(0, 47, 0.01)
h_a = rep(0, 4701)
for(i in 1:4701){
  v = spline_func(range_xa[i])
  h_a[i] = harzard_func(v)
}

harzard_ratio_a = as.data.frame(cbind(range_xa, h_a)) 
ggplot(data=harzard_ratio_a, aes(x=range_xa, y=h_a)) + geom_line(color='black', lwd=0.5) + ggtitle("harzard ratio of Arm_A") + 
xlab("The number of months") + ylab("Monthly deaths") +theme_bw()

```



In the same way as above, the case of Arm_B of the NCOG can also be obtained.
```{r}
## harzard function of Arm_B
num_B = nrow(Arm_B)
Km_B <- matrix(0, 76, 4)
copy_B = Arm_B
for(i in 1:76){
  Km_B[i,1] = nrow(copy_B)
  inter_B = copy_B %>% filter(t > 30.4*(i-1) & t <= 30.4*i)
  Km_B[i, 2] <- length(which(inter_B$d == 1))
  Km_B[i, 3] <- length(which(inter_B$d == 0))
  Km_B[i, 4] <- Km_B[i, 2] / Km_B[i,1]
  copy_B = copy_B %>% filter(t > 30.4*i)
}
x_b <- matrix(0, 76, 4)
for(i in 1:76){
  if(i <= 11){
    x_b[i,] <- c(1, i, (i-12)^2, (i-12)^3)
  }
  else{
    x_b[i,1:2] <- c(1, i)
  }
}
n_b = Km_B[,1]
y_b = Km_B[,2]
## Iteratively Rewighted least square(IRLS)
# initial value of alpha
alpha_0 = c(1, 0.01, 0.01, 0.01)
alpha = matrix(0, 1001, 4)
alpha[1,] = alpha_0
for(i in 1:1000){
  lambda = x_b %*% alpha[i,]
  
  # mu_k = n * exp(lambda_k)/ (1 + exp(lambda_k))
  mu = n_b * exp(lambda) / (1 + exp(lambda))
  
  # z_k = lambda_k + n / (mu_k * (n - mu_k)) * (y - mu_k)
  z = lambda  + (n_b/ (mu * (n_b - mu)) ) * (y_b - mu)
  
  
  # D_k = diag(mu_k(n - mu_k) / n)
  D = diag(as.numeric(mu*(n_b-mu)/n_b))
  
  
  # V_k = diag(mu_k (1 - mu_k/n))
  V = diag(as.numeric(mu* (1 - mu/n_b)))
  
  # W_k = D %*% solve(V) %*% D
  
  W = D
  
  # alpha_(k+1) = (X^t W_k X)^(-1) X^t W_k Z_k
  alpha[i+1, ] = solve((t(x_b) %*% W %*% x_b)) %*% t(x_b) %*% W %*% z
}
alpha_hat_b = alpha[1001,]

alpha_hat_b
```

Thus, the numerical mle of $\alpha$ in Arm_B case is $\hat{\alpha} = (-3.5586, -0.0219,  0.1262,  0.0132)^T$. Next, we'll draw the graph of harzard ratio.

```{r}
harzard_func <- function(x){
  h = 1/(1 + exp(-x%*%alpha_hat_b))
  return(h)
}
spline_func <- function(x){
  if(x<= 12){
    return(c(1, x, (x-12)^2,(x-12)^3))
  }
  else{
    return(c(1, x, 0, 0))
  }
}
range_xb = seq(0, 47, 0.01)
h_b = rep(0, 4701)
for(i in 1:4701){
  v = spline_func(range_xb[i])
  h_b[i] = harzard_func(v)
}
harzard_ratio_b = as.data.frame(cbind(range_xb, h_b)) 
ggplot(data=harzard_ratio_b, aes(x=range_xb, y=h_b)) + 
geom_line(color='red', lwd=0.5) +  ggtitle("harzard ratio of Arm_B") + xlab("The number of months") + 
ylab("Monthly deaths") + theme_bw()
```



So we can combine above two graphs.
```{r}
ggplot() + geom_line(data = harzard_ratio_a, aes(x=range_xa, y=h_a), color = "black", lwd = 0.5) +
  theme(legend.position = c(0.9,0.7)) +
  geom_line(data = harzard_ratio_b, aes(x = range_xb, y = h_b), color = 'red', lwd = 0.5) + 
  ggtitle("harzard ratio of Arm_A and Arm_B") +
  xlab("The number of months") + ylab("Monthly deaths") + 
  theme_bw()
```


## Problem 9.5
Why does the hypergeometric distribution enter into formula (9.24)?

### Solution
In the book, the notations mean $n_A$ = the number of at risk in Arm_A, $n_d$ = the number of deaths,
$n_B$ = the number of at risk in Arm_B, $n_s$ = the number of survived patients and y = the number of Arm_A deaths. 
The data of 2 by 2 display of month-6 for NCOG study are observed after the patients died which means we know $n_d$. Also, we know the number of patient using the treatments Arm_A and Arm_B respectively which means we know $n_A$ and $n_B$. Therefore, the sum of each columns and rows 2 by 2 contingency table is fixed. So y determines the other three table entires by substraction, so we are not losing any information by focusing on y. This is why the distribution of y follows the hypergeometric distribution.